Notes on conformal manifolds

### 2025-09-23 Draft_MLFlatDirections.tex
Update during meeting.

The AdS/CFT correspondence stands as one of the most profound dualities in theoretical physics, establishing a remarkable equivalence between string theory solutions on anti-de Sitter (AdS) spacetimes and conformal field theories (CFTs) living on their boundaries~\cite{Maldacena:1997re}. This correspondence has revolutionized our understanding of both quantum gravity and strongly coupled field theories, providing unprecedented insights into the holographic nature of gravity. Within this holographic framework, supergravity theories in AdS backgrounds serve as the low-energy effective descriptions of string theory compactifications, making them natural laboratories for exploring the gravitational side of the duality. The CFT side plays a crucial role in statistical physics, as CFT are known to describe phase transitions ; they are also fixed point of Renormalisation Group flow and this is this latter aspect that is our interest in this paper. An interesting question is indeed to know whether those CFT are isolated fixed points of RG flows, or belong to a continuous family of CFT. If this is the case, the space of continuous deformations that takes from one CFT to another is called a conformal manifold. These deformations are parametrised by exactly marginal operators, or in other words operators that preserves the conformal symmetry, \textit{i.e.} whose $\beta$ functions exactly vanish. %%ICI \ce{Needs some modifications.}

@@

	To address these challenges, we formulate the search for annihilating polynomials as a determination of a density probability on the space of polynomials that stresses higher probabilities on polynomials $p$ that minimises $\displaystyle \sum_{i=1}^{n_{\rm points}}p(\vec{x}^{(i)})^{2}$. Such a density probability is \textit{a priori} hard to find, but can be approximated using variations of Important Sampling methods~\cite{910e8e7d-2e70-317c-adf6-e21cd50755a1}. Importance sampling is a method to approximate expectation values with respect to a target distribution $d$ from a weighted average from another distribution $\pi$. It is based on the observation that the expectation
	\begin{equation}
		\mathbb{E}_{d}[f] = \int f(z)d(z)\ \d z
	\end{equation}
	of a function $f$ with respect to $d$ can be computed from the expectation $\mathbb{E}_{\pi}[w f]$, now computed with respect to the distribution $\pi$, with unnormalised $w(z) = d(z)/\pi(z)$. Therefore, if it is easier to draw a sample $\{z^{(k)}\}_{k\in\llbracket 1, n_{\rm particles}\rrbracket}$ from $\pi$ than from $d$, one can estimate $\mathbb{E}_{d}[f]$ from the weighted average
	\begin{equation}
		\mathbb{E}_{d}[f] \simeq \frac{\displaystyle \sum_{k=1}^{n_{\rm particles}} w(z^{(k)})\,f(z^{(k)})}{\displaystyle \sum_{i=1}^{n_{\rm particles}} w(z^{(k)})}.
	\end{equation}
	The accuracy of this estimation depends on the size of the sample and, crucially, on the choice of the distribution $\pi$.

	To have an accurate estimation of $d$, we employ an Annealed Importance Sampling (AIS) \cite{neal1998annealedimportancesampling} combined with Sequential Monte Carlo (SMC) methods \cite{del2006sequential,naesseth2024elementssequentialmontecarlo}. AIS aims at estimating $d$ sequentially from multiple intermediate distributions $\pi_{n}$ that smoothly transition from an initial, tractable distribution $d_{0}$ (chosen to be easy to sample from) to the complex target distribution $d$.
	These distributions are approximated by weighted samples $\{z^{(k)}_{n},w^{(k)}_{n}\}$ of particles, each representing a candidate symbolic expression, such that the resulting weighted average estimates $\pi_{n}$: \ce{Normalise the weights?}
	\begin{equation}
		\frac{\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}\,f(z^{(k)}_{n})}{\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}} \underset{n_{\rm particles}\rightarrow+\infty}{\longrightarrow} \mathbb{E}_{\pi_{n}}[f],
	\end{equation}
	or equivalently
	\begin{equation} \label{eq:pifromweights}
		\pi_{n}(z) \simeq \frac{\displaystyle \sum_{k=1}^{N} w^{(k)}_{n}\,\delta(z-z^{(k)}_{n})}{\displaystyle \sum_{k=1}^{N} w^{(k)}_{n}}.
	\end{equation}
	The $\pi_{n}$ are incrementally deformed using a Markov kernel to gradually reproduce the target $d$ and the particles get reweighed. This annealing process is guided by a temperature-like parameter that gradually emphasises the data likelihood, allowing for more efficient and tunable exploration of the polynomials landscape.

@@

	At each epoch $n$, $\pi_{n}$ is approximated \textit{via} a set of particles and weights $\{z_{n}^{(k)},\ w_{n}^{(k)}\}$ in the sense of eq.~\eqref{eq:pifromweights}. The sample of particles $z^{(k)}_{n}$ get mutated using a forward propagation Markov kernel $q(z^{(k)}_{n+1}\vert z^{(k)}_{n})$, defining the probability to get $z^{(k)}_{n+1}$ when the current state is $z^{(k)}_{n}$. For the implementation of the forward propagation kernel, we perform an AIS-style MCMC algorithm. 
	On every epoch, we make some move in the space of polynomials, and then accept or reject those new polynomials based on a fixed rate.
	The moves that have been allowed in our implementation are the following.
	\begin{itemize}
		\item \textbf{Coefficient pertubation}\quad Given a polynomial, we choose one of its coefficients at random and modify it by a Gaussian noise distributed as $\mathcal{N}(0,\sigma^2)$, \textit{e.g.}
		\begin{equation}
			2 x_1x_{2} + 3 x_2^2 \longmapsto  2.1 x_1x_{2} + 3 x_2^2\,.
		\end{equation}
		\item \textbf{Variable multiplication}\quad Given a polynomial, we pick randomly one of its monomials, and multiply it by one of the available variables, \textit{e.g.}
		\begin{equation}
			2 x_1x_{2} + 3 x_2^2 \longmapsto  2 x_1 x_2^{2} + 3 x_2^2\,.
		\end{equation}
		\item \textbf{Variable division}\quad Given a polynomial, we choose randomly one of its monomials, and divide it by one of its variables, \textit{e.g.}
		\begin{equation}
			2 x_1 x_2 + 3 x_2^2 \longmapsto  2 x_1 + 3 x_2^2\,.
		\end{equation}
	\end{itemize}
	Each of these operations is chosen randomly, with probabilities $p_{\rm shift}$, $p_{\rm multiply}$, and $p_{\rm divide}$.
	%		For the case of coefficient perturbations, the transition is symmetric. 
	%		However, this symmetry does not hold in general. For other types of operations—such as multiplication or division by a variable—the transition from \( z_{n-1}^k \) to \( z_n^k \) is generally not equivalent to the reverse transition from \( z_n^k \) to \( z_{n-1}^k \). In such cases $q(z_n^k \mid z_{n-1}^k) \ne q(z_{n-1}^k \mid z_n^k)$ and particular care must be taken in the computation of the backward propagation kernel.
	After performing these updates, the change is accepted based on the temperature-dependent rate
	\begin{equation}\label{eq:acceptanceratio}
		A(z_{n+1}^{(k)},z_{n}^{(k)}) = \mathrm{min} \left(1,\frac{ \gamma_{n+1}(z_{n+1}^{(k)}) q(z_{n}^{(k)} | z_{n+1}^{(k)}) }{ \gamma_{n}(z_{n}^{(k)}) q(z_{n+1}^{(k)} | z_{n}^{(k)}) }  \right)\,.
	\end{equation}
	We then draw $u\sim \mathrm{Uniform}(0,1)$, accept the new particle if $u<A(z_{n+1}^{(k)},z_{n}^{(k)})$ and reject it otherwise.
	The kernel $q(z_{n-1}^k | z_n^k)$ thus encapsulates the effects of performing these operations on the polynomials at step $n$. \ce{Why this comment?}

@@

	\subsection{Local analysis: extracting the dimension}
	Once the gradient descent has been completed and the flat directions sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we first perform some exploratory analyses to better understand the data. Specifically, we aim at determining the dimension of the manifold and whether it consists of a single connected component or multiple disjoint components (\textit{e.g.} two intersecting hyperplanes). To this end, we apply a local Principal Component Analysis (PCA,%\footnote{Principal Component Analysis is a dimensionality reduction technique that transforms high-dimensional data into a lower-dimensional space while preserving the maximum amount of variance. The method identifies orthogonal linear combinations of the original variables, called principal components, which are ordered by the proportion of total variance they explain. PCA operates by computing the eigenvalues and eigenvectors of the data covariance matrix, where the eigenvectors define the direction of maximum variance and the eigenvalues quantify the amount of variance captured along each direction. From a physicist point of view, as we are doing it locally, this gives access to information about the local tangent plane.}
see for example ref.~\cite{scikit-learn}). For each point, we identify its $k$ nearest neighbours and perform a PCA on that local neighbourhood. This procedure allows us to determine how many principal directions are needed to explain a given proportion $\epsilon$ of the data variance. In other words, it provides an estimate of the local dimensionality around each point, \textit{i.e.} the dimension of its local tangent space. We perform this analysis for several values of $k$, namely $k \in \{5, 10, 20, 50, 100\}$, and we fix $\epsilon = 0.99$. The results are presented in fig.~\ref{fig:local_pca}. We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned dimension 2. This can be attributed to the fact that if the genuine dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality for a fraction of the points. Additionally, for $k \geq 20$, we observe an increasing number of points being assigned dimensions 4 or 5. This behaviour can be explained by the loss of locality when the number of neighbours becomes too large: increasing $k$ results in a coarser approximation, and the algorithm may then incorporate points that are no longer truly local. This artificial enlargement of the neighbourhood can cause the estimated local dimensionality to rise. We thus infer that the manifold under investigation has an intrinsic dimension of 3.

@@

	After determining the local dimension of the space of solutions, we need to ascertain its topology. 
One possible scenario is that our data consists of several three-dimensional manifolds, intersecting at least at the origin $\vec{X}=0$, and the points previously identified with dimension $4$ may lie at the intersections of these manifolds. 
Consider for example the intersection of two lines: at the intersection point, the local dimension estimated by the previous PCA algorithm would be 2. 
To rule out this possibility, we apply a clustering algorithm on the points with tangent spaces of dimension 3 only, as identified by the PCA with $k=20$. 
This, way, we remove the possible intersection points with local dimension 4 or 5. 
For the purpose of the clustering, we use the density-based algorithm HDBSCAN~\cite{10.1007/978-3-642-37456-2_14}. 
As it can be observed in fig.~\ref{triangular_plot_1_2_4_8_10}, the points obtained after the gradient descent are denser around the origin, leading the algorithm to disregard points with a norm greater than 1. 
To avoid this effect, we de-sample the densest areas by randomly selecting no more than 10 points per 0.25-sided hyper-cube, and apply the HDBSCAN algorithm on the remaining points, with minimum cluster size set to 5 points. 
For our initial $10^5$ points, the algorithm identifies one cluster made of 5 points, one other with 71,993 points, and fails to assign any cluster to 218 points. We show in fig.~\ref{fig:3dplots} some 3d projections of the data to visualize the clustering. In these scatter plots, orange points belong to the main cluster, while blue points are those that the algorithm failed to assign to any cluster. The smallest cluster, made of 5 points, is likely an artefact of local fluctuations in the data density and is not interpreted as physically meaningful. The size of the different points has been adjusted to facilitate the visualisation. From visual inspection, it appears that the unassigned (blue) points lie mostly on the boundary of the sampled region. We therefore interpret their unassigned status not as evidence of belonging to another manifold, but rather as a result of insufficient local density near the edges of the dataset. Thus, the algorithm does indicate that over 99\% of the data belongs to a single dominant cluster.

@@

Regarding the temperature schedule, as can be seen in \eqref{eq:gamman}, what matters to determine $\gamma_{n}$ is not the inverse temperature, but rather the product of the inverse temperature and the loss function. Therefore, instead of implementing an arbitrary schedule for $\beta$, we employed an adaptive temperature approach, whose aim is to ensure that the number of particles to be updated at each run is close to
\begin{equation} \label{eq:eqadaptivetemperatureschedule}
	\texttt{adaptive\_temperature\_ratio} = 0.8 - 0.5 \times \left(\frac{\texttt{n}}{\texttt{n\_epochs}}\right).
\end{equation}
In other words, if the number of accepted particles after the MCMC step at a given \texttt{n} is higher than the target~\eqref{eq:eqadaptivetemperatureschedule}, we decrease the temperature and hence increase $\beta_{n}$, and \textit{vice versa} if this number is lower than \texttt{adaptive\_temperature\_ratio}.

This encourages exploration during the early stages of training and gradually transitions to a more selective regime.

>>>>

The AdS/CFT correspondence stands as one of the most profound dualities in theoretical physics, establishing a remarkable equivalence between string theory solutions on anti-de Sitter (AdS) spacetimes and conformal field theories (CFTs) living on their boundaries~\cite{Maldacena:1997re}. This correspondence has revolutionized our understanding of both quantum gravity and strongly coupled field theories, providing unprecedented insights into the holographic nature of gravity. Within this holographic framework, supergravity theories in AdS backgrounds serve as the low-energy effective descriptions of string theory compactifications, making them natural laboratories for exploring the gravitational side of the duality. The CFT side plays a crucial role in statistical physics, as CFT are known to describe phase transitions ; they are also fixed point of Renormalisation Group flow and this is this latter aspect that is our interest in this paper. An interesting question is indeed to know whether those CFT are isolated fixed points of RG flows, or belong to a continuous family of CFT. If this is the case, the space of continuous deformations that takes from one CFT to another is called a conformal manifold. These deformations are parametrised by exactly marginal operators, or in other words operators that preserves the conformal symmetry, \textit{i.e.} whose $\beta$ functions exactly vanish. ICI \ce{Needs some modifications.}

@@

	To address these challenges, we formulate the search for annihilating polynomials as a determination of a density probability on the space of polynomials that stresses higher probabilities on polynomials $p$ that minimises $\displaystyle \sum_{i=1}^{n_{\rm points}}p(\vec{x}^{(i)})^{2}$. Such a probability density is \textit{a priori} hard to find, but can be approximated using variations of Important Sampling methods~\cite{910e8e7d-2e70-317c-adf6-e21cd50755a1}. Importance sampling is a method to parametrise expectation values with respect to a target distribution $d$, from a weighted average using another distribution $\pi$. It is based on the observation that the expectation
	\begin{equation}
		\mathbb{E}_{d}[f] = \int f(z)d(z)\ \d z
	\end{equation}
	of a function $f$ with respect to $d$ can be computed from the expectation $\mathbb{E}_{\pi}[w f]$, now computed with respect to the distribution $\pi$, with unnormalised $w(z) = d(z)/\pi(z)$. Therefore, if it is easier to draw a sample $\{z^{(k)}\}_{k\in\llbracket 1, n_{\rm particles}\rrbracket}$ from $\pi$ than from $d$, one can estimate $\mathbb{E}_{d}[f]$ from the weighted average
	\begin{equation}
		\mathbb{E}_{d}[f] \simeq \frac{\displaystyle \sum_{k=1}^{n_{\rm particles}} w(z^{(k)})\,f(z^{(k)})}{\displaystyle \sum_{i=1}^{n_{\rm particles}} w(z^{(k)})}.
	\end{equation}
	The accuracy of this estimation depends on the size of the sample and, crucially, on the choice of the distribution $\pi$. \ce{Add conclusion saying that you can parametrise $d$ using a sample and weights, and eq. as (4.4).}

	To have an accurate estimation of $d$, we employ an Annealed Importance Sampling (AIS) \cite{neal1998annealedimportancesampling} combined with Sequential Monte Carlo (SMC) methods \cite{del2006sequential,naesseth2024elementssequentialmontecarlo}. AIS aims at estimating $d$ sequentially from multiple intermediate distributions $\pi_{n}$ that smoothly transition from an initial, tractable distribution $d_{0}$ (chosen to be easy to sample from) to the complex target distribution $d$.
	These distributions are approximated by weighted samples $\{z^{(k)}_{n},w^{(k)}_{n}\}$ of particles, each representing a candidate symbolic expression, such that the resulting weighted average estimates $\pi_{n}$: \ce{Normalise the weights?}
	\begin{equation}
		\frac{\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}\,f(z^{(k)}_{n})}{\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}} \underset{n_{\rm particles}\rightarrow+\infty}{\longrightarrow} \mathbb{E}_{\pi_{n}}[f],
	\end{equation}
	or equivalently
	\begin{equation} \label{eq:pifromweights}
		\pi_{n}(z) \simeq \frac{\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}\,\delta(z-z^{(k)}_{n})}{\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}}.
	\end{equation}
	The $\pi_{n}$ are incrementally deformed using a Markov kernel to gradually reproduce the target $d$ and the particles get reweighed. This annealing process is guided by a temperature-like parameter that gradually emphasises the data likelihood, allowing for more efficient and tunable exploration of the polynomials landscape.

@@

	At each epoch $n$, $\pi_{n}$ is approximated \textit{via} a set of particles and weights $\{z_{n}^{(k)},\ w_{n}^{(k)}\}$ in the sense of eq.~\eqref{eq:pifromweights}. The sample of particles $z^{(k)}_{n}$ get mutated using a forward propagation Markov kernel $q(z^{(k)}_{n+1}\vert z^{(k)}_{n})$, defining the probability to get $z^{(k)}_{n+1}$ when the current state is $z^{(k)}_{n}$. For the implementation of the forward propagation kernel, we perform an AIS-style MCMC algorithm. 
	On every epoch, we make some move in the space of polynomials, and then accept or reject those new polynomials based on a fixed rate.
	The moves that have been allowed in our implementation are the following.
	\begin{itemize}
		\item \textbf{Coefficient perturbation}\quad Given a polynomial, we choose one of its coefficients at random and modify it by a Gaussian noise distributed as $\mathcal{N}(0,\sigma^2)$, \textit{e.g.}
		\begin{equation}
			2 x_1x_{2} + 3 x_2^2 \longmapsto  2.1 x_1x_{2} + 3 x_2^2\,.
		\end{equation}
		\item \textbf{Variable multiplication}\quad Given a polynomial, we pick randomly one of its monomials, and multiply it by one of the available variables, \textit{e.g.}
		\begin{equation}
			2 x_1x_{2} + 3 x_2^2 \longmapsto  2 x_1 x_2^{2} + 3 x_2^2\,.
		\end{equation}
		\item \textbf{Variable division}\quad Given a polynomial, we choose randomly one of its monomials, and divide it by one of its variables, \textit{e.g.}
		\begin{equation}
			2 x_1 x_2 + 3 x_2^2 \longmapsto  2 x_1 + 3 x_2^2\,.
		\end{equation}
	\end{itemize}
	Each of these operations is chosen randomly, with probabilities $p_{\rm shift}$, $p_{\rm multiply}$, and $p_{\rm divide}$.
	%		For the case of coefficient perturbations, the transition is symmetric. 
	%		However, this symmetry does not hold in general. For other types of operations—such as multiplication or division by a variable—the transition from \( z_{n-1}^k \) to \( z_n^k \) is generally not equivalent to the reverse transition from \( z_n^k \) to \( z_{n-1}^k \). In such cases $q(z_n^k \mid z_{n-1}^k) \ne q(z_{n-1}^k \mid z_n^k)$ and particular care must be taken in the computation of the backward propagation kernel.
	After performing these updates, the change is accepted based on the temperature-dependent rate
	\begin{equation}\label{eq:acceptanceratio}
		A(z_{n+1}^{(k)},z_{n}^{(k)}) = \mathrm{min} \left(1,\frac{ \gamma_{n+1}(z_{n+1}^{(k)}) q(z_{n}^{(k)} | z_{n+1}^{(k)}) }{ \gamma_{n}(z_{n}^{(k)}) q(z_{n+1}^{(k)} | z_{n}^{(k)}) }  \right)\,.
	\end{equation}
	We then draw $u\sim \mathrm{Uniform}(0,1)$, accept the new particle if $u<A(z_{n+1}^{(k)},z_{n}^{(k)})$ and reject it otherwise.

@@

	\subsection{Local analysis: extracting the dimension}
	Once the gradient descent has been completed and the flat directions sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we first perform some exploratory analyses to better understand the data. Specifically, we aim at determining the dimension of the manifold and whether it consists of a single connected component or multiple disjoint components (\textit{e.g.} two intersecting hyperplanes). To this end, we apply a local Principal Component Analysis (PCA).\footnote{Principal Component Analysis is a dimensionality reduction technique that transforms high-dimensional data into a lower-dimensional space while preserving the maximum amount of variance, see for example ref.~\cite{scikit-learn}.} % The method identifies orthogonal linear combinations of the original variables, called principal components, which are ordered by the proportion of total variance they explain. PCA operates by computing the eigenvalues and eigenvectors of the data covariance matrix, where the eigenvectors define the direction of maximum variance and the eigenvalues quantify the amount of variance captured along each direction. From a physicist point of view, as we are doing it locally, this gives access to information about the local tangent plane.}
	For each point, we identify its $k$ nearest neighbours and perform a PCA on that local neighbourhood. This procedure allows us to determine how many principal directions are needed to explain a given proportion $\epsilon$ of the data variance. In other words, it provides an estimate of the local dimensionality around each point, \textit{i.e.} the dimension of its local tangent space. We perform this analysis for several values of $k$, namely $k \in \{5, 10, 20, 50, 100\}$, and we fix $\epsilon = 0.99$. The results are presented in fig.~\ref{fig:local_pca}. We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned dimension 2. This can be attributed to the fact that if the genuine dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality for a fraction of the points. Additionally, for $k \geq 20$, we observe an increasing number of points being assigned dimensions 4 or 5. This behaviour can be explained by the loss of locality when the number of neighbours becomes too large: increasing $k$ results in a coarser approximation, and the algorithm may then incorporate points that are no longer truly local. This artificial enlargement of the neighbourhood can cause the estimated local dimensionality to rise. We thus infer that the manifold under investigation has an intrinsic dimension of 3.

@@

	After determining the local dimension of the space of solutions, we need to ascertain its topology. 
One possible scenario is that our data consists of several three-dimensional manifolds, intersecting at least at the origin $\vec{X}=0$, and the points previously identified with dimension $4$ may lie at the intersections of these manifolds. 
Consider for example the intersection of two lines: at the intersection point, the local dimension estimated by the previous PCA algorithm would be 2. 
To rule out this possibility, we apply a clustering algorithm on the points with tangent spaces of dimension 3 only, as identified by the PCA with $k=20$. 
This, way, we remove the possible intersection points with local dimension 4 or 5. 
For the purpose of the clustering, we use the density-based algorithm HDBSCAN~\cite{10.1007/978-3-642-37456-2_14}. 
As it can be observed in fig.~\ref{triangular_plot_1_2_4_8_10}, the points obtained after the gradient descent are denser around the origin, leading the algorithm to disregard points with a norm greater than 1. 
To avoid this effect, we de-sample the densest areas by randomly selecting no more than 10 points per 0.25-sided hyper-cube, and apply the HDBSCAN algorithm on the remaining 72,216 points, with minimum cluster size set to 5 points. 
The algorithm identifies one cluster made of 5 points, one other with 71,993 points, and fails to assign any cluster to 218 points. We show in fig.~\ref{fig:3dplots} some 3d projections of the data to visualize the clustering. In these scatter plots, orange points belong to the main cluster, while blue points are those that the algorithm failed to assign to any cluster. The smallest cluster, made of 5 points, is likely an artefact of local fluctuations in the data density and is not interpreted as physically meaningful. The size of the different points has been adjusted to facilitate the visualisation. From visual inspection, it appears that the unassigned (blue) points lie mostly on the boundary of the sampled region. We therefore interpret their unassigned status not as evidence of belonging to another manifold, but rather as a result of insufficient local density near the edges of the dataset. Thus, the algorithm does indicate that over 99\% of the data belongs to a single dominant cluster.

@@

Regarding the temperature schedule, as can be seen in \eqref{eq:gamman}, what matters to determine $\gamma_{n}$ is not the inverse temperature, but rather the product of the inverse temperature and the loss function. Therefore, instead of implementing an arbitrary schedule for $\beta$, we employed an adaptive temperature approach, whose aim is to ensure that the number of particles to be updated at each run is close to
\begin{equation} \label{eq:eqadaptivetemperatureschedule}
	\texttt{adaptive\_temperature\_ratio} = 0.8 - 0.5 \times \left(\frac{\texttt{n}}{\texttt{n\_epochs}}\right).
\end{equation}
In other words, if the number of accepted particles after the MCMC step at a given \texttt{n} is higher than the target~\eqref{eq:eqadaptivetemperatureschedule}, we decrease the temperature and hence increase $\beta_{n}$, and \textit{vice versa} if this number is lower than \texttt{adaptive\_temperature\_ratio}. This encourages exploration during the early stages of training and gradually transitions to a more selective regime.
